---
title: "CABG Readmission Analysis"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
library(shiny)
library(flexdashboard)
library(ggplot2)
library(DT)
library(dplyr)
library(plotly)

# Load Original Data
original_dataset <- read.csv("simulated_cabg_data.csv") %>% select(-readmission_status)

# Reactive values for data storage
data_store <- reactiveVal(original_dataset)

# Function to update dataset
update_data <- function(new_data) {
  current_data <- data_store()
  updated_data <- bind_rows(current_data, new_data)
  data_store(updated_data)
}

# Variable labels
var_labels <- c(
  "albumin" = "Albumin (g/dL)",
  "INR" = "INR",
  "last_hematocrit" = "Last Hematocrit (%)",
  "total_bilirubin" = "Total Bilirubin (mg/dL)",
  "ejection_fraction" = "Ejection Fraction (%)",
  "last_creatinine_level" = "Last Creatinine (mg/dL)",
  "last_wbc_count" = "Last WBC Count (cells/μL)",
  "last_a1c_level" = "Last A1C Level (%)",
  "platelet_count" = "Platelet Count (PLT/μL)",
  "meld_score" = "MELD Score"
)
```

# Dashboard Overview

## Dashboard Summary

Coronary Artery Bypass Grafting (CABG) is a common surgical procedure worldwide. Predicting readmission risk can improve patient care, support clinical decision-making, and reduce healthcare costs. This dashboard explores the top 10 biological factors associated with 30-day readmission after CABG.The data set is simulated based on the distribution of the original. The model is a simple logistic regression and requires further refinement therefore results are not conclusive and don't provide any useful medical information. The main goal is to demonstrate proficiency in building dashboards and this layout can be used with proper model development to get robust results.  

In this dashboard users can:

- Download dataset
- Explore dataset summaries and distributions of key variables.
- Add new patient observations.
- Predict readmission risk using a logistic regression model.
- Visualize the contribution of each predictor to the calculated risk

#  Data Overview

## Access Data and Exploration

```{r}
fluidRow(
  
  # Left side: Download and Preview Dataset
  column(5,
         h3("Download Dataset"),
         downloadButton("download_dataset", "Download Data"),
         br(), br(),  # Spacer
         h3("Dataset Preview"),
         DTOutput("table_display")
  ),
  
  # Right side: Explore Variable Distributions
  column(7,
         h3("Explore Variable Distributions"),
         selectInput("eda_var", "Select Variable to Explore", choices = c(
           "Albumin (g/dL)" = "albumin",
           "INR" = "INR",
           "Last Hematocrit (%)" = "last_hematocrit",
           "Total Bilirubin (mg/dL)" = "total_bilirubin",
           "Ejection Fraction (%)" = "ejection_fraction",
           "Last Creatinine Level (mg/dL)" = "last_creatinine_level",
           "Last WBC Count (cells/μL)" = "last_wbc_count",
           "Last A1C Level (%)" = "last_a1c_level",
           "Platelet Count (PLT/μL)" = "platelet_count",
           "MELD Score" = "meld_score"
         )),
         plotlyOutput("variable_distribution_plot")
  )
)


```



```{r}
# Render the Data Table
output$table_display <- renderDT({
  datatable(data_store(), options = list(scrollX = TRUE)) 
})

# Download Handler
output$download_dataset <- downloadHandler(
  filename = function() {
    paste0("CABG_dataset_", Sys.Date(), ".csv")
  },
  content = function(file) {
    write.csv(data_store(), file, row.names = FALSE)
  }
)

# Render Plotly Distribution
output$variable_distribution_plot <- renderPlotly({
  req(input$eda_var) 
  selected_var <- input$eda_var
  readable_var <- var_labels[selected_var]
  
  ggplot(data_store(), aes_string(x = selected_var, fill = "as.factor(readmission_binary)")) +
    geom_density(alpha = 0.5) +
    labs(
      title = paste("Distribution of", readable_var, "by Readmission Status"),
      x = readable_var,
      fill = "Readmission"
    ) +
    theme_minimal() +
    theme(plot.title = element_text(size = 10))
})


```



### Key Takeaway

The red (readmission=1) indicated patiens who have been readmitted within 30 days where are the green(readmission=0) indicates those who have not been readmitted. From the density graphs, patients who were readmitted, on average, had higher albumin, hematocrit, meld score,  and lower last creatinine level, WBC count, A1C levels.



# Adding New Observations and Risk Prediction

## Add New Observation

```{r}
fluidRow(
  column(4, numericInput("new_albumin", "Albumin (g/dL)", value = 2, min = 1.1, max = 3.9)),
  column(4, numericInput("new_INR", "INR", value = 1, min = 0.27, max = 12.4)),
  column(4, numericInput("new_last_hematocrit", "Last Hematocrit (%)", value = 40, min = 10, max = 87))
  )


fluidRow(
  column(4, numericInput("new_total_bilirubin", "Total Bilirubin (mg/dL)", value = 0.5, min = 0.1, max = 6)),
  column(4, numericInput("new_ejection_fraction", "Ejection Fraction (%)", value = 35, min = 10, max = 83)),
  column(4, numericInput("new_last_creatinine_level", "Last Creatinine (mg/dL)", value = 0.5, min = 0.3, max = 18.3))
)

fluidRow(
  column(4, numericInput("new_last_wbc_count", "Last WBC Count (cells/μL)", value = 7, min = 2, max = 20)),
  column(4, numericInput("new_last_a1c_level", "Last A1C (%)", value = 6, min = 0.9, max = 18)),
  column(4, numericInput("new_platelet_count", "Platelet Count (PLT/μL)", value = 27000, min = 25300, max = 857000))
)

fluidRow(
  column(4, numericInput("new_meld_score", "MELD Score", value = 9, min = 6.4, max = 37.8)),
  column(4, selectInput("new_readmission_binary", "Readmission Status", choices = c("No" = 0, "Yes" = 1)))
)

actionButton("add_data", "Add Observation")

observeEvent(input$add_data, {
  new_entry <- data.frame(
    albumin = input$new_albumin,
    INR = input$new_INR,
    last_hematocrit = input$new_last_hematocrit,
    total_bilirubin = input$new_total_bilirubin,
    ejection_fraction = input$new_ejection_fraction,
    last_creatinine_level = input$new_last_creatinine_level,
    last_wbc_count = input$new_last_wbc_count,
    last_a1c_level = input$new_last_a1c_level,
    platelet_count = input$new_platelet_count,
    meld_score = input$new_meld_score,
    readmission_binary = as.numeric(input$new_readmission_binary)
  )
  update_data(new_entry)
})

```



## Risk Prediction

```{r}
#developing logistic model
logit_model <- reactive({
  glm(readmission_binary ~ ., data = data_store(), family = binomial)
})
predict_risk <- function(input_data) {
  input_data <- input_data %>% mutate(across(everything(), as.numeric)) # Ensure numeric format
  predict(logit_model(), newdata = input_data, type = "response")
}

predict_input <- reactive({
  data.frame(
    albumin = input$new_albumin,
    INR = input$new_INR,
    last_hematocrit = input$new_last_hematocrit,
    total_bilirubin = input$new_total_bilirubin,
    ejection_fraction = input$new_ejection_fraction,
    last_creatinine_level = input$new_last_creatinine_level,
    last_wbc_count = input$new_last_wbc_count,
    last_a1c_level = input$new_last_a1c_level,
    platelet_count = input$new_platelet_count,
    meld_score = input$new_meld_score
  )
})

#show the predicted risk 
renderUI({
  HTML(paste0("<h2>Predicted Readmission Risk: ", round(predict_risk(predict_input()) * 100, 2), "%</h2>"))
})


```

### Variable contribution to risk
```{r}
# visualizing predictor contribution to calculated risk
output$contrib_plot <- renderPlotly({
  
  coef_values <- coef(logit_model())[-1]  # Remove intercept
  contribs <- coef_values * unlist(predict_input())

  contribs_df <- data.frame(
    Variable = names(coef_values),
    Contribution = contribs
  )

    # Replace raw variable names with readable labels
  contribs_df$Variable <- var_labels[contribs_df$Variable]
  p <- ggplot(contribs_df, aes(x = reorder(Variable, Contribution), y = Contribution, fill = Contribution)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Predictor's Contribution to Readmission Risk",
         x = "Feature", y = "Contribution") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Ensure full label visibility

    theme(plot.title = element_text(size = 9))
  ggplotly(p)
})

plotlyOutput("contrib_plot")

```

### Key TakeAway:

Generally, A1C level, WBC count, last creatinine, ejection fraction, bilirubin contribute to higher predicted risk of readmission. In contrast, the rest have negative association with risk of readmission.

